Title
========================================================

## Part I
How many observations and how many variables in the diamonds dataset?

How many order factors in the variables?

```{r}
library(ggplot2)
data(diamonds)
str(diamonds)
```

what letters represents the best diamond?

```{r}
levels(diamonds$color)
```

## Part II
Create a histogram of the price of all the diamonds in the diamond data set.
```{r}
library(ggthemes) 
theme_set(theme_minimal(24)) 
qplot(x = price, data = diamonds)
```

## Part III
Describe the shape and center of the price distribution. Include summary statistics  like mean and median

```{r}
summary(diamonds$price)
```

Long tail distributed

## Part IV
How many diamonds cost less than $500,250 or more than 15000
```{r}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >= 15000)
```


## Part V
Explore the largest peak in the price histogram you created earlier.

Try limiting the x-axis, altering the bin width, and setting different breaks on the x-axis.

```{r}
qplot(x = price, data = diamonds)
qplot(x = price, data = diamonds, xlim = c(0,2500), binwidth = 100)
qplot(x = price, data = diamonds, xlim = c(0,2500), binwidth = 200)
qplot(x = price, data = diamonds, xlim = c(0,2500), binwidth = 500)
ggsave('priceHistogram.png')
```

## Part VI
Which cut has the highest priced dimond?
```{r}
qplot(x = price, data = diamonds) + 
  facet_wrap(~cut)

diamonds$cut[which.max(diamonds$price)]
diamonds$cut[which.min(diamonds$price)]
by(diamonds$price, diamonds$cut, summary)
```

## Part VII

add a parameter to facet_wrap so that the y-axis in the histograms is not fixed. You want the y-axis to be different for each histogram.
```{r}
qplot(x = price, data = diamonds) + 
  facet_wrap(~cut, scales = "free")
```

## Part VIII

Create a histogram of price per carat and facet it by cut. 
```{r}
qplot(x = log10(price/carat+1), data = diamonds, binwidth = 0.3)+
  facet_wrap(~cut, scales = "free")
```

## Part IX
Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color.

```{r}
by(diamonds$price, diamonds$cut, summary)
by(diamonds$price, diamonds$clarity, summary)
by(diamonds$price, diamonds$color, summary)
qplot(x = cut, y = price, data = diamonds, geom = 'boxplot')
qplot(x = clarity, y = price, data = diamonds, geom = 'boxplot')
qplot(x = color, y = price, data = diamonds, geom = 'boxplot')
```


## Part X
what is the price range for the middle 50% of diamonds with color D?
what is the price range for the middle 50% of diamonds with color J?
what is the IQR for the diamonds with best color?
what is the IQR for the diamonds with worst color?
```{r}
summary(diamonds$price[diamonds$color == "D"])
summary(diamonds$price[diamonds$color == "J"])
IQR(diamonds$price[diamonds$color == "D"])
IQR(diamonds$price[diamonds$color == "J"])
```


## Part XI
Investigate the price per carat of diamonds across the different colors of diamonds using boxplots.

```{r}
qplot(x = color, y = price/carat, data = diamonds, geom = 'boxplot')+
    coord_cartesian(ylim = c(0,6000))
```


## Part XII
investigate the weight of the diamonds using frequency polygon.
what carat size has a count greater than 2000?
```{r}
qplot(x = carat, data = diamonds,binwidth = 0.1,
      geom = "freqpoly")+
  scale_x_continuous(limits = c(0,6), breaks = seq(0,6,0.1))

table(round(diamonds$carat,1))
```


## Part XIII

Which month contains the most number of birthdays?
How many birthdays are in each month?
Which day of the year has the most number of birthdays?
```{r}
friends <- read.csv('birthdaysExample.csv', stringsAsFactors = F)
friends$dates <- as.Date(friends$dates, format = "%m/%d/%y")
friends$months <- months(friends$dates)
qplot(months, data = friends)
sort(table(friends$months),decreasing = T)[1]
sort(table(friends$dates),decreasing = T)[1]
```

