Lesson 5
========================================================

```{r}
library(ggplot2)
library(ggthemes)
library(gridExtra)
theme_set(theme_solarized())
pf <- read.delim('pseudo_facebook.tsv')
```

## Third Qualitative Variable
```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot()

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + 
  geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
  geom_line(aes(color = gender),stat = 'summary', fun.y = median)
```


```{r}
library(dplyr)
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age,gender) %>%
  summarise( mean_friend_count = mean(friend_count),
             median_friend_count = median(friend_count),
             n = n()) %>%
  ungroup() %>%
  arrange(age)
head(pf.fc_by_age_gender)
```

### Plotting Conditional Summaries
```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count), data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender))
```

### Reshaping Data
```{r}
library(reshape2)
head(pf.fc_by_age_gender)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')
pf.fc_by_age_gender.wide$ratio <- pf.fc_by_age_gender.wide$female/pf.fc_by_age_gender.wide$male
head(pf.fc_by_age_gender.wide)
```


```{r}
pf.fc_by_age_gender.wide <- pf.fc_by_age_gender %>% 
  group_by(age) %>% 
  summarise(male = median_friend_count[gender == 'male'], 
                      female = median_friend_count[gender == 'female'], 
                      ratio = female / male) %>% 
  arrange(age) 

head(pf.fc_by_age_gender.wide)
```

### Ratio Plot
```{r Ratio Plot}

# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

# This assignment is not graded and
# will be marked as correct when you submit.
ggplot(aes(x = age, y = ratio), data = pf.fc_by_age_gender.wide) +
  geom_line()+
  geom_hline(yintercept = 1 ,alpha = 0.3, linetype = 2)
```

### Third Quantitative Variable
```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

### Cut a Variable
```{r Cut a Variable}
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]


pf$year_joined.bucket <- cut(pf$year_joined, breaks = c(2004,2009,2011,2012,2014))
```

### Plotting it All Together
```{r Plotting it All Together}
# Create a line graph of friend_count vs. age
# so that each year_joined.bucket is a line
# tracking the median user friend_count across
# age. This means you should have four different
# lines on your plot.
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary', fun.y = median)
```

### Plot the Grand Mean
```{r Plot the Grand Mean}
# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash
ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary', fun.y = mean)+
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

### Friending Rate

```{r Friending Rate}
# what's median friend rate
# what's max friend rate
with(subset(pf, tenure >= 1), summary(friend_count/tenure))
```

### Friendships Initiated
```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.
ggplot(aes(x = tenure, y = friendships_initiated/tenure),
       data = subset(pf, tenure >= 1))+ 
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary', fun.y = mean)
```

### Bias-Variance Tradeoff Revisited
Notes:
```{r Bias-Variance Tradeoff Revisited}
p1 <- ggplot(aes(x = tenure, y = friendships_initiated/tenure),
       data = subset(pf, tenure >= 1))+ 
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary', fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)
grid.arrange(p1,p2,p3,p4, ncol = 1)
```

```{r, echo = FALSE}
# Instead of geom_line(), use geom_smooth() to add a smoother to the plot.
# You can use the defaults for geom_smooth() but do color the line
# by year_joined.bucket
ggplot(aes(x = tenure, y = friendships_initiated/tenure),
       data = subset(pf, tenure >= 1))+ 
  geom_smooth(aes(color = year_joined.bucket))
```

### Histograms Revisited
```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)
yo$id <- as.factor(yo$id)
ggplot(aes(x = price), data =yo) +
  geom_histogram()
```


### Number of Purchases
```{r Number of Purchases}
yo <- transform(yo, all.purchase = strawberry + blueberry + pina.colada + plain + mixed.berry)
```

### Prices over Time
```{r Prices over Time}
ggplot(aes(x = all.purchase, fill = 'orange'), data = yo) +
  geom_histogram(binwidth = 1)

ggplot(aes(x = time, y = price), data = yo) +
  geom_jitter(alpha = 1/10)
```

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(4230)
sample.ids <- sample(levels(yo$id),16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids))+
  facet_wrap(~ id)+
  geom_line()+
  geom_point(aes(size = all.purchase), pch = 1)
```

### Scatterplot Matrix
```{r}
#install.packages('GGally')
library(GGally)
set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset),1000),])
```

### Heat Maps
```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

