Lesson 4
========================================================
read in data
```{r, echo = FALSE}
library(ggplot2)
library(ggthemes)
theme_set(theme_economist())
pf <- read.csv("pseudo_facebook.tsv", sep = "\t")
```

### Scatterplots and Perceived Audience Size
qplot
```{r}
qplot(x = age, y = friend_count, data = pf)
```

equalvent ggplot
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point()
```

limit the x axis
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point() +
  xlim(13,90)
```

add transparent and jitter to look through the overlapping points
### Overplotting
```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) +
  xlim(13,90)
```

#### What do you notice in the plot?
Response: most young user have friends less than 1000, not significantly bigger than older user.

### Coord_trans()
```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0)) +
  xlim(13,90) +
  coord_trans(y = "sqrt")
```

```{r}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h=0))+
  xlim(13,90) + 
  coord_trans(y = "sqrt")
```

### Conditional Means

```{r Conditional Means}
#install.packages('dplyr')
library(dplyr)
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups, 
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age,age)
```

equal to
```{r}
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

pf.fc_by_age
```

# Plot mean friend count vs. age using a line graph.
# Be sure you use the correct variable names
# and the correct data frame. You should be working
# with the new data frame created from the dplyr
# functions. The data frame is called 'pf.fc_by_age'.

# Use geom_line() rather than geom_point to create
# the plot. You can look up the documentation for
# geom_line() to see what it does.

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age)+
  geom_line()
```

### Overlaying Summaries with Raw Data
```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20,
             position = position_jitter(h = 0),
             color = "orange") +
  xlim(13,90) +
  coord_trans(y = "sqrt") +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = quantile, probs = .1,
            linetype = 2, color = 'blue') + 
  geom_line(stat = "summary", fun.y = quantile, probs = .9,
            linetype = 2, color = 'red') +
  geom_line(stat = "summary", fun.y = quantile, probs = .5,
            linetype = 2, color = 'green')
```

Better plot
```{r}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20,
             position = position_jitter(h = 0),
             color = "orange") +
  coord_cartesian(xlim = c(13,70), ylim = c(0,1000)) +
  geom_line(stat = "summary", fun.y = mean) +
  geom_line(stat = "summary", fun.y = quantile, probs = .1,
            linetype = 2, color = 'blue') + 
  geom_line(stat = "summary", fun.y = quantile, probs = .9,
            linetype = 2, color = 'red') +
  geom_line(stat = "summary", fun.y = quantile, probs = .5,
            linetype = 2, color = 'green')
```

### Correlation
```{r Correlation}
names(cor.test(pf$friend_count, pf$age))
round(cor.test(pf$friend_count, pf$age, meathod = "pearson")$estimate,3)
```

equal
```{r}
with(pf, cor.test(age, friend_count,method = "pearson"))
```

### Correlation on Subsets
```{r Correlation on Subsets}
with(subset(pf, pf$age <= 70) , cor.test(age, friend_count))
```

### Correlation Methods
```{r}
with(subset(pf, pf$age <= 70), cor.test(age, friend_count,method = "spearman"))
```

## Create Scatterplots
# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x). Use any of the
# techniques that you've learned so far to
# modify the plot.
```{r}
ggplot(aes(x = www_likes_received,  y = likes_received), data = pf) +
  geom_point()
```

### Strong Correlations
```{r Strong Correlations}
ggplot(aes(x = www_likes_received,  y = likes_received), data = pf) +
  geom_point(alpha = 1/4) +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95))+
  geom_smooth(method = "lm", color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
round(with(pf, cor.test(www_likes_received, likes_received))$estimate,3)
```

### More Caution with Correlation
```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data(Mitchell)
```

```{r Temp vs Month}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor(Mitchell$Month, Mitchell$Temp)
```

### Making Sense of Data
```{r Making Sense of Data}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()+
  scale_x_discrete(breaks = seq(0,203,12))
```

### A New Perspective
```{r  New Perspective, fig.width = 10, fig.height = 3}
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()+
  scale_x_discrete(breaks = seq(0,203,12))
```

### Understanding Noise: Age to Age Months
```{r Age to Age Months}
ggplot(aes(x = age, y = friend_count_mean), data = pf.fc_by_age) +
  geom_line()

head(pf.fc_by_age)
pf.fc_by_age[17:19,]
```

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf$age + (12 - pf$dob_month) / 12
```

Programming Assignment
```{r Programming Assignment}
age_with_months_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months <- summarise(age_with_months_groups,
                                 friend_count_mean = mean(friend_count),
                                 friend_count_median = median(friend_count),
                                 n = n())
arrange(pf.fc_by_age_months, age_with_months)
```

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, pf.fc_by_age_months$age_with_months <= 71)) +
  geom_line()
```

### Smoothing Conditional Means
```{r}
q1 <- ggplot(aes(x = age, y = friend_count_mean), 
            data = subset(pf.fc_by_age, pf.fc_by_age$age <= 71)) +
              geom_line() +
  geom_smooth()

q2 <- ggplot(aes(x = age_with_months, y = friend_count_mean), 
       data = subset(pf.fc_by_age_months, pf.fc_by_age_months$age_with_months <= 71)) +
  geom_line() + 
  geom_smooth()

q3 <- ggplot(aes(x = round(age/5)*5, y = friend_count),
             data = subset(pf, pf$age <= 71)) +
  geom_line(stat = 'summary', fun.y = mean)

library(gridExtra)
grid.arrange(q1,q2,q3,ncol = 1)
```

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

