Lesson 3
========================================================
### Pseudo-Facebook User Data
```{r Pseudo-Facebook User Data}
getwd()
pf <- read.csv('pseudo_facebook.tsv', sep = "\t")
pf <- read.delim('pseudo_facebook.tsv')
?read.csv
?read.delim
names(pf)
```

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
install.packages('ggplot2')
library(ggplot2)
qplot(x = dob_day, data = pf)
qplot(x = dob_day, data = pf)+
  scale_x_discrete(breaks = 1:31)
```

***

Run the following code in R to get other themes. 
```{r}
install.packages('ggthemes', dependencies = TRUE) 
library(ggthemes) 
theme_set(theme_minimal(24)) 
```

Plot using ggplot 
```{r}
ggplot(aes(x = dob_day), data = pf) + 
  geom_histogram() + 
  scale_x_discrete(breaks = 1:31)
```


#### What are some things that you notice about this histogram?

Response: huge number of people have birthday the first day of months 

### Faceting
histogram for each month

```{r Faceting}
qplot(x = dob_day, data = pf)+
  scale_x_discrete(breaks = 1:31)+
  facet_wrap(~dob_month, ncol = 4)
```

equalent but hard to look at using facet_grid
```{r}
qplot(x = dob_day, data = pf)+
  scale_x_discrete(breaks = 1:31)+
  facet_grid(. ~ dob_month)
```


ggplot equalient
```{r}
ggplot(data = pf, aes(x = dob_day)) + 
  geom_histogram() + 
  scale_x_discrete(breaks = 1:31) + 
  facet_wrap(~dob_month)
```

#### Let us take another look at our plot. What stands out to you here?
Response: January 1st
  
### Friend Count

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(x = friend_count, data = pf)
```

#### How is this plot similar to Moira's first plot?
Response: long tail

### Limiting the Axes
```{r Limiting the Axes}
qplot(x = friend_count, data = pf, xlim = c(0,1000))
```


```{r}
qplot(x = friend_count, data = pf)+
  scale_x_continuous(limits = c(0,1000))
```

### Adjusting the Bin Width

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```

```{r}
qplot(x = friend_count, data = pf, binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))+
  facet_wrap(~gender)
```


### Omitting NA Values
```{r Omitting NA Values}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))+
  facet_wrap(~gender)
```

```{r}
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))+
  facet_wrap(~gender)
```


### Statistics 'by' Gender

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count,pf$gender,summary)
```

#### Who on average has more friends: men or women?
Response: WOMEN

#### What's the difference between the median friend count for women and men?
Response: 77

### Tenure

```{r Tenure}
qplot(x = tenure, data =pf, binwidth = 30,
      color = I('black'), fill = I('#099009'))
```


#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x = tenure/365, data =pf, binwidth = 1,
      color = I('black'), fill = I('#758381'))
```

### Labeling Plots
```{r Labeling Plots}
qplot(x = tenure/365, data = pf,
      xlab = "Number of years using Facebook",
      ylab = "Nuber of users in sample",
      color = I('black'), fill = I('#F79420'))+
  scale_x_continuous(breaks = seq(1,7,1), lim = c(0,7))
```

### User Ages

```{r User Ages}
qplot(x = age, data = pf, binwidth = 1,
      color = I('black'), fill = I('#5760AB'))

qplot(x = age, data = pf, binwidth = 1,
      fill = I('#5760AB')) +
  scale_x_continuous(breaks = seq(0,113,5))
```

### Transforming Data
```{r}
summary(pf$friend_count)
summary(log10(pf$friend_count))
summary(log10(pf$friend_count+1))
summary(sqrt(pf$friend_count))
```


```{r}
install.packages('gridExtra') 
library(gridExtra) 
```

```{r}
p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = log10(friend_count+1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf)
grid.arrange(p1,p2,p3,ncol = 1)
```

```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()
grid.arrange(p1,p2,p3,ncol = 1)
```


### Add a Scaling Layer
```{r Add a Scaling Layer}
q1 <- qplot(x = log10(friend_count), data = pf)
q2 <- ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()+
  scale_x_log10()
grid.arrange(q1,q2, ncol = 2)
```

### Frequency Polygons
```{r Frequency Polygons}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10,
      geom = "freqpoly", color = gender)+
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```


```{r}
qplot(x = friend_count, y = ..count../sum(..count..), 
      data = subset(pf, !is.na(gender)), binwidth = 10,
      geom = "freqpoly", color = gender) +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50))
```


```{r}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) + 
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')
```

### Likes on the Web
```{r}
by(pf$www_likes, pf$gender, sum)
```



```{r Likes on the Web}
qplot(x = www_likes, data = subset(pf,!is.na(gender)), 
      geom = "freqpoly", color = gender) +
  scale_x_continuous() +
  scale_x_log10()

ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) + 
  geom_freqpoly(aes(color = gender)) + 
  scale_x_log10()
```

### Box Plots

```{r Box Plots}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), 
      geom = "boxplot")+
  scale_y_continuous(limits = c(0, 1000)) 

qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)), 
      geom = "boxplot", ylim = c(0,1000))
```

Correct way!
```{r}
qplot(x = gender, y = friend_count, data = subset(pf, !is.na(gender)),
      geom = "boxplot") +
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```

#### On average, who initiated more friendships in our sample: men or women?
```{r Friend Requests by Gender}
qplot(x = gender, y = friendships_initiated, data = subset(pf, !is.na(gender)), geom = 'boxplot') +
  coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

### Getting Logical
```{r Getting Logical}
summary(pf$mobile_likes)
summary(pf$mobile_likes > 0 )

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```


Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!