Lesson 6
========================================================

### Scatterplot Review

```{r Scatterplot Review}
library(ggplot2)
library(ggthemes)
theme_set(theme_economist())
qplot(data = diamonds, x = carat, y = price,
      xlim = c(0, quantile(diamonds$carat, 0.99)),
      ylim = c(0, quantile(diamonds$price, 0.99)))+
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21)
```

### Price and Carat Relationship
Non linear relationship and the variance increases when carat increase.
```{r}
qplot(data = diamonds, x = carat, y = price,
      xlim = c(0, quantile(diamonds$carat, 0.99)),
      ylim = c(0, quantile(diamonds$price, 0.99)))+
  stat_smooth(method = 'lm') +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21, alpha = 1/4)
```


### ggpairs Function
```{r ggpairs Function}
# install these if necessary
# install.packages('memisc')
# install.packages('lattice')
# install.packages('MASS')
# install.packages('car')
# install.packages('reshape')
# install.packages('plyr')

# load the ggplot graphics package and the others
library(GGally)
library(scales)
library(memisc)


# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, params = c(shape = I('.'), outlier.shape = I('.')))
```

What are some things you notice in the ggpairs output?
Response:
clarity and color are related to price.
carat and x,y,z are correlated with price.

### The Demand of Diamonds
```{r The Demand of Diamonds}
# Create two histograms of the price variable
# and place them side by side on one output image.

# We¡¯ve put some code below to get you started.

# The first plot should be a histogram of price
# and the second plot should transform
# the price variable using log10.

# Set appropriate bin widths for each plot.
# ggtitle() will add a title to each histogram.
library(gridExtra)
plot1 <- qplot(price, data = diamonds, binwidth = 100) + 
  ggtitle('Price')
plot2 <- qplot(log10(price+1), data = diamonds, binwidth = 0.01) +
  ggtitle('Price (log10)')
grid.arrange(plot1, plot2, ncol = 2)
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function(){
  trans_new('cuberoot',
            transform = function(x) {x^(1/3)},
            inverse = function(x){ x^3})
  }
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
# Add a layer to adjust the features of the
# scatterplot. Set the transparency to one half,
# the size to three-fourths, and jitter the points.
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 1/2, size = 3/4, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

### Price vs. Carat and Clarity
Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
#install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

### Price vs. Carat and Cut
Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

### Cut and Price
```{r}
table(diamonds$cut)
```

### Price vs. Carat and Color
Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = F,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

### Building the Linear Model
```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

### A Bigger, Better Data Set
```{r A Bigger, Better Data Set}
library('bitops')
load('BigDiamonds.Rda')
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set

```{r Building a Model Using the Big Diamonds Data Set}
diamondsbig$logprice <- log(diamondsbig$price)
m1 <- lm(logprice ~ I(carat^(1/3)),
         data = diamondsbig[diamondsbig$price < 10000 &
                              diamondsbig$cert == "GIA",])
m2 <- update(m1, ~. + carat)
m3 <- update(m2, ~. + cut)
m4 <- update(m3, ~. + color)
m5 <- update(m4, ~. + clarity)
mtable(m1,m2,m3,m4,m5)
```

## Predictions
Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601
```{r}
#Be sure you have loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

